---
title: "Data Engine Primary Request Function"
subtitle: "request function definition"
author: "Conor Flynn"
date: "04/18/2023"
output:
  pdf_document: default
  html_document:
    toc: true
    number_sections: true
    df_print: paged
---
# Start by loading the proper libraries:
```{r setup, include=FALSE}
library(knitr)
library(plyr)
library(dplyr)
library(jsonlite)
library(stringr)
library(tidyr)
```

## Define request function:

# DESCRIPTION:
# This function serves as the primary request point for making internal requests to the defi data engine. Note that should a request be
# made outside the function, please ensure the necessary format depicted within it is used otherwise an internal error may occur.
# Note that you MUST be connected to the RPI network (such that a connection can be made to defi-de.idea.rpi.edu) otherwise this
# function will not work.

# INPUT:
#               - protocol:     The name of the protocol to receive data from.
# [optional]    - properties:   All properties used in the call (note typically REST API URL parameters).
# [optional]    - headers:      All headers used in the call (note typically REST API URL headers).
# [optional]    - startdate:    Starting date to retrieve data from. In format 'yyyy-MM-dd' i.e. 2023-04-01
# [optional]    - enddate:      Ending date to retrieve data from (non-inclusive). In format 'yyyy-MM-dd' i.e. 2023-04-01

# OUTPUT:
# list containing two elements:
#               - $response:    Contains all response information as listed below:
#                 - $response:  Value denoting status of call TO the engine. Code 200 denotes connection was received by engine properly. 
#                 - $code:      Code returned by engine based on internal schema. Full list of codes can be found here
#                                 (https://github.rpi.edu/DataINCITE/IDEA-DeFi-CRAFT/wiki/Response-Codes)
#                 - $message:   Message response accompanying code should the response be irregular.
#                 - $data:      Data returned by call should any be requested.
#               - $df:          Data frame containing all data parsed from the call.
```{r}
request <- function(protocol, properties = "", headers = "", startdate = "", enddate = "") {
  suppressWarnings({
    #Create socket and get destination which tells the engine where to put the data
    socket <- socketConnection("defi-de.idea.rpi.edu", 61200, blocking=TRUE)
    destination <- readLines(socket, 1)
    
    formatted_properties = ""
    if(properties != "")
      formatted_properties = paste("properties", "&&&", properties, "&&&")
    
    formatted_headers = ""
    if(headers != "")
      formatted_headers = paste("headers", "&&&", headers, "&&&")
    
    formatted_startdate = ""
    formatted_enddate = ""
    if(startdate != "" && enddate != "") {
      formatted_startdate = paste("start_date", "&&&", startdate, "&&&")
      formatted_enddate = paste("end_date", "&&&", enddate, "&&&")
    }
    
    #Build the request delimited by &&&
    #Similar to a GET request in the way we handle parameters
    request.raw <- paste(
    				"SRC", "&&&", "RQST", "&&&",
    				"type", "&&&", protocol, "&&&",
    				formatted_properties,
    				formatted_headers,
    				formatted_startdate,
    				formatted_enddate,
    				"destination", "&&&", destination, "&&&",
    				"\n", sep="")
    
    # remove all spaces from request
    request.data <- str_replace_all(request.raw, " ", "")
    
    #Write this request back to the socket to tell engine what we want
    writeLines(request.data, socket)
    
    # define a list which will store the individual rows
    rows = list()
    
    #Now the engine will begin feeding us data
    #We grab the first to initialize the data var and then we continue listening\
    counter <- 0
    response <- ""
    while (TRUE) {
      temp <- readLines(socket, 1)
      
      if(temp == '') 
      {
        print("Read empty string. Check engine logs or refresh configuration.")
        next
      }
      
      # if line is heartbeat then acknowledge and continue
      if (grepl("<<<heartbeat>>>", temp, fixed=TRUE))
      {
        print(paste0("Heartbeat read for ", protocol))
        next
      }
      
      # if line is response then process and terminate
      else if (grepl("<<<response>>>", temp, fixed=TRUE)) 
      {
        temp <- readLines(socket, 1)
        while(grepl("<<<heartbeat>>>", temp, fixed=TRUE)) {
          temp <- readLines(socket, 1)
        }
        response <- fromJSON(temp)
        break
      }
      
      # increment processed line counter
      counter <- counter + 1
      if(counter %% 1000 == 0){
        print(paste0("Processed ", counter, " lines for ", protocol))
      }
      
      # add data point line to data frame
      tryCatch(expr={
        rows[[counter]] = data.frame(fromJSON(temp))
      },
        error=function(e){
        message("Heartbeat exception caught and not parsed.")    
      })
    }
    
    output <- list("response"=response, "df"=do.call(rbind.fill, rows))
    close(socket)
    return(output)
  })
}
```

## Sample function call to the protocol graph-aave-borrows
```{r}
startdate <- "2022-01-01"
enddate <- "2022-01-04"

# submit a request
graph.borrows <- request("graph-aave-borrows", "", "", startdate, enddate)
# for each request validate the code was successful
# if not the return the response with an empty dataframe
if(graph.borrows$response$code != 200)
  return(graph.borrows)
```