---
title: "GetTransactions"
author: "Conor Flynn"
date: "2023-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Define necessary libraries
```{r}
library(knitr)
library(plyr)
library(dplyr)
library(jsonlite)
library(stringr)
library(tidyr)
library(randomNames)
```

## Define request function:

# NOTE:
# This function is officially defined in DataEnginePrimaryFunctions.Rmd and does not need to be redefined
# for every use-case. Here it is defined for and because there are requests made.

# DESCRIPTION:
# This function serves as the primary request point for making internal requests to the defi data engine. Note that should a request be
# made outside the function, please ensure the necessary format depicted within it is used otherwise an internal error may occur.
# Note that you MUST be connected to the RPI network (such that a connection can be made to defi-de.idea.rpi.edu) otherwise this
# function will not work.

# INPUT:
#               - protocol:     The name of the protocol to receive data from.
# [optional]    - properties:   All properties used in the call (note typically REST API URL parameters).
# [optional]    - headers:      All headers used in the call (note typically REST API URL headers).
# [optional]    - startdate:    Starting date to retrieve data from. In format 'yyyy-MM-dd' i.e. 2023-04-01
# [optional]    - enddate:      Ending date to retrieve data from (non-inclusive). In format 'yyyy-MM-dd' i.e. 2023-04-01

# OUTPUT:
# list containing two elements:
#               - $response:    Contains all response information as listed below:
#                 - $response:  Value denoting status of call TO the engine. Code 200 denotes connection was received by engine properly. 
#                 - $code:      Code returned by engine based on internal schema. Full list of codes can be found here (https://github.rpi.edu/DataINCITE/IDEA-DeFi-CRAFT/wiki/Response-Codes)
#                 - $message:   Message response accompanying code should the response be irregular.
#                 - $data:      Data returned by call should any be requested.
```{r}
request <- function(protocol, properties = "", headers = "", startdate = "", enddate = "") {
  suppressWarnings({
    #Create socket and get destination which tells the engine where to put the data
    socket <- socketConnection("defi-de.idea.rpi.edu", 61200, blocking=TRUE)
    destination <- readLines(socket, 1)
    
    formatted_properties = ""
    if(properties != "")
      formatted_properties = paste("properties", "&&&", properties, "&&&")
    
    formatted_headers = ""
    if(headers != "")
      formatted_headers = paste("headers", "&&&", headers, "&&&")
    
    formatted_startdate = ""
    formatted_enddate = ""
    if(startdate != "" && enddate != "") {
      formatted_startdate = paste("start_date", "&&&", startdate, "&&&")
      formatted_enddate = paste("end_date", "&&&", enddate, "&&&")
    }
    
    #Build the request delimited by &&&
    #Similar to a GET request in the way we handle parameters
    request.raw <- paste(
    				"SRC", "&&&", "RQST", "&&&",
    				"type", "&&&", protocol, "&&&",
    				formatted_properties,
    				formatted_headers,
    				formatted_startdate,
    				formatted_enddate,
    				"destination", "&&&", destination, "&&&",
    				"\n", sep="")
    
    # remove all spaces from request
    request.data <- str_replace_all(request.raw, " ", "")
    
    #Write this request back to the socket to tell engine what we want
    writeLines(request.data, socket)
    
    # define a list which will store the individual rows
    rows = list()
    
    #Now the engine will begin feeding us data
    #We grab the first to initialize the data var and then we continue listening\
    counter <- 0
    response <- ""
    while (TRUE) {
      temp <- readLines(socket, 1)
      
      if(temp == '') 
      {
        print("Read empty string. Check engine logs or refresh configuration.")
        next
      }
      
      # if line is heartbeat then acknowledge and continue
      if (grepl("<<<heartbeat>>>", temp, fixed=TRUE))
      {
        print(paste0("Heartbeat read for ", protocol))
        next
      }
      
      # if line is response then process and terminate
      else if (grepl("<<<response>>>", temp, fixed=TRUE)) 
      {
        temp <- readLines(socket, 1)
        while(grepl("<<<heartbeat>>>", temp, fixed=TRUE)) {
          temp <- readLines(socket, 1)
        }
        response <- fromJSON(temp)
        break
      }
      
      # increment processed line counter
      counter <- counter + 1
      if(counter %% 1000 == 0){
        print(paste0("Processed ", counter, " lines for ", protocol))
      }
      
      # add data point line to data frame
      tryCatch(expr={
        rows[[counter]] = data.frame(fromJSON(temp))
      },
        error=function(e){
        message("Heartbeat exception caught and not parsed.")    
      })
    }
    
    output <- list("response"=response, "df"=do.call(rbind.fill, rows))
    close(socket)
    return(output)
  })
}
```


## Define get_users function:

# DESCRIPTION:
# This function loads all user data from The Graph using the request() function defined above.
# Only one call needs to be made to this function so long as the data is loaded in the cache.
# Since it is a process-intensive call, it is recommended to only call this once and then
# pass the output as a parameter to the get_data() function.

# INPUT:

# OUTPUT:
# list containing two elements:
#               - $response:    Contains all response information as listed below:
#                 - $response:  Value denoting status of call TO the engine. Code 200 denotes connection was received by engine properly. 
#                 - $code:      Code returned by engine based on internal schema. Full list of codes can be found here
#                                 (https://github.rpi.edu/DataINCITE/IDEA-DeFi-CRAFT/wiki/Response-Codes)
#                 - $message:   Message response accompanying code should the response be irregular.
#                 - $data:      Data returned by call should any be requested.
#               - $df:          Data frame containing data returned from 'graph-aave-users' request.
```{r}
get_users <- function() {
  # retrieve users and then parse with aliases
  graph.users <- request("graph-aave-users", "", "")
  if(graph.users$response$code != 200)
    return(graph.users)
  
  # define data frame
  graph.users.df <- graph.users$df
  
  aliases = NULL
  set.seed(69420)
  
  while(length(aliases[,1]) < length(graph.users.df$id)){
    alias <- randomNames(1000, name.order = "first.last", name.sep = " ", sample.with.replacement = FALSE)
    aliases <- aliases %>%
      bind_rows(data.frame(alias)) %>%
      distinct()
  }
  
  aliases <- aliases %>%
    head(length(graph.users.df$id))
  
  userAliases <- bind_cols(graph.users, aliases) %>%
    mutate(version = "V2",
           deployment = "Mainnet")
  
 output <- list("response"=graph.users$response, "df"=userAliases)
}
```


## Define get_data function:

# DESCRIPTION:
# This function will request all necessary data to compute the transaction dataframe.
# This includes multiple calls to The Graph which can be seen below.

# INPUT:
#               - startdate:    Starting date to retrieve data from. In format 'yyyy-MM-dd' i.e. 2023-04-01
#               - enddate:      Ending date to retrieve data from (non-inclusive). In format 'yyyy-MM-dd' i.e. 2023-04-01
#               - users:        Response from get_users() function that contains all user data. If not passed then it will
#                                 request it in this function.

# OUTPUT:
# Dataframe containing all returned data.
```{r}
get_data <- function(startdate, enddate, users=NULL) {
  ### retrieve all data from defi engine:
  
  # submit a request
  graph.borrows <- request("graph-aave-borrows", "", "", startdate, enddate)
  # for each request validate the code was successful
  # if not the return the response with an empty dataframe
  if(graph.borrows$response$code != 200)
    return(graph.borrows)
  
  # repeat for all following requests
  graph.collaterals <- request("graph-aave-collaterals", "", "", startdate, enddate)
  if(graph.collaterals$response$code != 200)
    return(graph.collaterals)
  
  graph.deposits <- request("graph-aave-deposits", "", "", startdate, enddate)
  if(graph.deposits$response$code != 200)
    return(graph.deposits)
  
  graph.flashloans <- request("graph-aave-flash-loans", "", "", startdate, enddate)
  if(graph.flashloans$response$code != 200)
    return(graph.flashloans)
  
  graph.liquidations <- request("graph-aave-liquidations", "", "", startdate, enddate)
  if(graph.liquidations$response$code != 200)
    return(graph.liquidations)
  
  graph.pricehist <- request("graph-aave-price-history-items", "", "", startdate, enddate)
  if(graph.pricehist$response$code != 200)
    return(graph.pricehist)
  
  graph.redeems <- request("graph-aave-redeems", "", "", startdate, enddate)
  if(graph.redeems$response$code != 200)
    return(graph.redeems)
  
  graph.repays <- request("graph-aave-repays", "", "", startdate, enddate)
  if(graph.repays$response$code != 200)
    return(graph.repays)
  
  graph.reservehist <- request("graph-aave-reserve-params-hist-items", "", "", startdate, enddate)
  if(graph.reservehist$response$code != 200)
    return(graph.reservehist)
  
  graph.reserves <- request("graph-aave-reserves", "", "")
  if(graph.reserves$response$code != 200)
    return(graph.reserves)
  
  graph.swaps <- request("graph-aave-swaps", "", "", startdate, enddate)
  if(graph.swaps$response$code != 200)
    return(graph.swaps)
  
  graph.userreserves <- request("graph-aave-user-reserves", "", "", startdate, enddate)
  if(graph.userreserves$response$code != 200)
    return(graph.userreserves)
  
  # retrieve users and then parse with aliases
  graph.users <- users
  if(is.null(users)) { 
    # retrieve users and then parse with aliases
    graph.users <- request("graph-aave-users", "", "")
    if(graph.users$response$code != 200)
      return(graph.users)
    
    # define data frame
    graph.users.df <- graph.users$df
    
    aliases = NULL
    set.seed(69420)
    
    while(length(aliases[,1]) < length(graph.users.df$id)){
      alias <- randomNames(1000, name.order = "first.last", name.sep = " ", sample.with.replacement = FALSE)
      aliases <- aliases %>%
        bind_rows(data.frame(alias)) %>%
        distinct()
    }
    
    aliases <- aliases %>%
      head(length(graph.users.df$id))
    
    userAliases <- bind_cols(graph.users, aliases) %>%
      mutate(version = "V2",
             deployment = "Mainnet")
    
    graph.users <- list("response"=graph.users$response, "df"=userAliases)
  }
  
  # format into output list
  output <- list(borrows=graph.borrows, collaterals=graph.collaterals, deposits=graph.deposits,
                 flashloans=graph.flashloans, liquidations=graph.liquidations, pricehist=graph.pricehist,
                 redeems=graph.redeems, repays=graph.repays, reservehist=graph.reservehist, reserves=graph.reserves,
                 swaps=graph.swaps, userreserves=graph.userreserves, users=graph.users)
}
```

## Define get_transactions function:

# DESCRIPTION:
# This function takes input returned from the get_data(startdate, enddate, users) function to parse a table with
# all properly formatted transaction data.

# INPUT:
#               - data:       Response from the get_data() function containing all properly formatted data.

# OUTPUT:
# Dataframe containing all formatted transaction data.
```{r}
# get transactions function includes calls to both AmberData, GraphQL, and DeFiLlama
get_transactions <- function(data) {
  ### parse all data into proper raw naming conventions
  
  # load raw transaction tables:
  rawBorrows <- data$borrows$df
  rawCollaterals <- data$collaterals$df
  rawDeposits <- data$deposits$df
  rawLiquidations <- data$liquidations$df
  rawRedeems <- data$redeems$df
  rawRepays <- data$repays$df
  rawSwaps <- data$swaps$df
  rawFlashLoans <- data$flashloans$df
  
  # load raw reserve information:
  rawReserveInfo <- data$reserves$df
  rawReserveParamsHistory <- data$reservehist$df
  
  # select aliases
  aliases <- data$users$df %>%
    select(id, alias)
  
  # select reserve info
  reserveInfo <- rawReserveInfo %>%
    select(id,
       symbol,
       decimals)
  
  # select reserve parameter history
  reserveParamsHistory <- rawReserveParamsHistory %>%
    mutate(txID = str_sub(id, start = 1, end = 66)) %>%
    left_join(reserveInfo, by = c("reserve" = "id"))
  
  # mutate and parse borrows df
  borrows <- rawBorrows %>%
    mutate(type = "borrow",
           id = str_extract(id, "0x\\w+")) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(amount = as.numeric(amount) / (10^as.numeric(decimals))) %>%
    mutate(amountUSD = as.numeric(amount) * as.numeric(priceInUsd)) %>%
    mutate(priceInEth = as.numeric(priceInEth) / (10^18)) %>%
    mutate(amountETH = as.numeric(amount) * as.numeric(priceInEth)) %>%
    mutate(reserve = symbol) %>%
    mutate(user = user) %>%
    mutate(onBehalfOf = caller) %>%
    mutate(borrowRate = as.numeric(borrowRate) / (10^25)) %>%
    mutate(pool = pool) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    left_join(aliases, by = c("onBehalfOf" = "id")) %>%
    dplyr::rename(onBehalfOfAlias = alias) %>%
    select(id, type, timestamp, user, userAlias, onBehalfOf, onBehalfOfAlias, pool, 
           reserve, amount, amountUSD, amountETH, borrowRate, borrowRateMode) %>%
    drop_na() %>%
    distinct()
  
  # mutate and parse collaterals df
  collaterals <- rawCollaterals %>%
    mutate(type = "collateral",
           id = str_extract(id, "0x\\w+")) %>%
    mutate(user = user, pool = pool) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(reserve = symbol) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    select(id, timestamp, user, userAlias, pool, reserve, fromState, toState, type) %>%
    drop_na() %>%
    distinct()
  
  # mutate and parse deposits df
  deposits <- rawDeposits %>%
    mutate(type = "deposit",
           id = str_extract(id, "0x\\w+")) %>%
    mutate(user = user, pool = pool) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(reserve = symbol) %>%
    mutate(onBehalfOf = caller) %>%
    mutate(amount = as.numeric(amount) / (10^as.numeric(decimals))) %>%
    mutate(amountUSD = as.numeric(amount) * as.numeric(priceInUsd)) %>%
    mutate(priceInEth = as.numeric(priceInEth) / (10^18)) %>%
    mutate(amountETH = as.numeric(amount) * as.numeric(priceInEth)) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    left_join(aliases, by = c("onBehalfOf" = "id")) %>%
    dplyr::rename(onBehalfOfAlias = alias) %>%
    select(id, timestamp, type, amount, amountUSD, amountETH, reserve, user, userAlias, 
           onBehalfOf, onBehalfOfAlias, pool) %>%
    drop_na() %>%
    distinct()
  
  # mutate and parse liquidations df
  liquidations <- rawLiquidations %>%
    mutate(type = "liquidation",
           id = str_extract(id, "0x\\w+"),
           user = user, pool = pool) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "principalReserve" = "reserve", "timestamp")) %>%
    mutate(principalAmount = as.numeric(principalAmount) / 10^as.numeric(decimals),
           principalAmountUSD = as.numeric(priceInUsd) * as.numeric(principalAmount),
           principalAmountETH = as.numeric(priceInEth) * as.numeric(principalAmount) / 10^18) %>%
    select(id, timestamp, type, user, liquidator, pool, 
           principalAmount, principalReserve = symbol, principalAmountUSD, principalAmountETH, 
           collateralAmount, collateralReserve) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "collateralReserve" = "reserve", "timestamp")) %>%
    mutate(collateralAmount = as.numeric(collateralAmount) / 10^as.numeric(decimals),
           collateralAmountUSD = as.numeric(priceInUsd) * as.numeric(collateralAmount),
           collateralAmountETH = as.numeric(priceInEth) * as.numeric(collateralAmount) / 10^18) %>%
    select(id, timestamp, type, user, liquidator, pool,
           principalAmount, principalReserve, principalAmountUSD, principalAmountETH,
           collateralAmount, collateralReserve = symbol, collateralAmountUSD, collateralAmountETH) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    left_join(aliases, by = c("liquidator" = "id")) %>%
    dplyr::rename(liquidatorAlias = alias) %>%
    distinct()
  
  # mutate and parse redeems df
  redeems <- rawRedeems %>%
    mutate(type = "redeem",
           id = str_extract(id, "0x\\w+")) %>%
    mutate(user = user, pool = pool) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(reserve = symbol) %>%
    mutate(onBehalfOf = to) %>%
    mutate(amount = as.numeric(amount) / (10^as.numeric(decimals))) %>%
    mutate(amountUSD = as.numeric(amount) * as.numeric(priceInUsd)) %>%
    mutate(priceInEth = as.numeric(priceInEth) / (10^18)) %>%
    mutate(amountETH = as.numeric(amount) * as.numeric(priceInEth)) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    left_join(aliases, by = c("onBehalfOf" = "id")) %>%
    dplyr::rename(onBehalfOfAlias = alias) %>%
    select(id,timestamp, type, amount, amountUSD, amountETH, reserve, user, userAlias, 
          onBehalfOf, onBehalfOfAlias, priceInUsd, pool) %>%
    drop_na() %>%
    distinct()
  
  # mutate and parse repays df
  repays <- rawRepays %>%
    mutate(type = "repay",
           id = str_extract(id, "0x\\w+"))%>%
    mutate(user = user, pool = pool) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(reserve = symbol) %>%
    mutate(onBehalfOf = repayer) %>%
    mutate(amount = as.numeric(amount) / (10^as.numeric(decimals))) %>%
    mutate(amountUSD = as.numeric(amount) * as.numeric(priceInUsd)) %>%
    mutate(priceInEth = as.numeric(priceInEth) / (10^18)) %>%
    mutate(amountETH = as.numeric(amount) * as.numeric(priceInEth)) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    left_join(aliases, by = c("onBehalfOf" = "id")) %>%
    dplyr::rename(onBehalfOfAlias = alias) %>%
    select(id, timestamp, type, amount, amountUSD, amountETH, reserve, user, userAlias, 
           onBehalfOf, onBehalfOfAlias, priceInUsd, pool) %>%
    drop_na() %>%
    distinct()
  
  # mutate and parse swaps df
  swaps <- rawSwaps %>%
    mutate(type = "swap",
           id = str_extract(id, "0x\\w+")) %>%
    mutate(user = user, pool = pool) %>%
    mutate(stableBorrowRate = as.numeric(stableBorrowRate) / (10^25), variableBorrowRate = as.numeric(variableBorrowRate) / (10^25)) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(reserve = symbol) %>%
    left_join(aliases, by = c("user" = "id")) %>%
    dplyr::rename(userAlias = alias) %>%
    select(id, timestamp, type, reserve, user, userAlias, pool, borrowRateModeTo, 
           borrowRateModeFrom, stableBorrowRate = stableBorrowRate.x, variableBorrowRate = variableBorrowRate.x) %>%
    drop_na() %>%
    distinct()
  
  # mutate and parse flashloans df
  flashLoans <- rawFlashLoans %>%
    mutate(type = "flashLoan",
           id = str_extract(id, "0x\\w+")) %>%
    left_join(reserveParamsHistory, by = c("id" = "txID", "timestamp", "reserve")) %>%
    mutate(reserve = symbol,
           amount = as.numeric(amount)) %>%
    select(id, timestamp, type, reserve, target, pool, amount, totalFee, target) %>%
    drop_na() %>%
    distinct()
  
  # define binded clean transactions df
  cleanedTransactions <- borrows %>%
    bind_rows(collaterals) %>%
    bind_rows(deposits) %>%
    bind_rows(liquidations) %>%
    bind_rows(redeems) %>%
    bind_rows(repays) %>%
    bind_rows(swaps) %>%
    bind_rows(flashLoans)
  
  # return successful output with given successful response code
  output <- list("response"=list("code"=200, "data"="", "response"=200, "message"="Successful Response"), "df"=cleanedTransactions)
  output
}
```


## Sample Code:

# First we make a request to get user data:
```{r}
users <- get_users()
```

# Next we write a call to get_data to get all needed data:
```{r}
startdate <-  "2022-01-01"
enddate <-    "2022-01-10"
data <-       get_data(startdate, enddate, users)
```

# Finally make a call to get all transactional data:
```{r}
transactions <- get_transactions(data)
```